---
title: "Project 2"
author: "Justin Feathers & Luke Perkins"
date: "2022-10-08"
output:
  github_document:
    toc: true
    toc_depth: 2
    df_print: tibble
---

The purpose of this vignette is to explain how to use a function created for
interacting with a specific API, as well as demonstrate various types of
analyses that can be conducted on the data obtained from its use. 
The [Beer API](https://www.openbrewerydb.org/) is a free API that 
contains publicly available address and contact information for breweries, 
cideries, brewpubs, and bottleshops all over the world.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(httr)
library(tidyverse)
library(jsonlite)
```

# Overview

The following packages were used to create this document:  

  * `knitr`: used to document code in R Markdown format  
  * `httr`: used to connect to the API  
  * `tidyverse`: used for pipe operators and plotting  
  * `jsonlite`: used for interacting with JSON in the API  


## Requirements

The following packages were used to create this document: 

  + `knitr`: used to document code in R Markdown format  
  + `httr`: used to connect to the API  
  + `tidyverse`: used for pipe operators and plotting  
  + `jsonlite`: used for interacting with JSON in the API  

# API Function

The function we will use to query the API uses one endpoint, or connection point
between the API and server, at <https://api.openbrewerydb.org/breweries>. Six
different modifications, each passed as separate arguments of the function, 
allow the user to filter the returned API data based on any combination of 
single values of the variables of interest.


```{r}
get_brewery <- function(city = NULL, state = NULL, usa = TRUE, type = NULL, 
                        name = NULL, postal = NULL, latitude = NULL, 
                        longitude = NULL){
  url <- "https://api.openbrewerydb.org/breweries?per_page=50"

  if (!is.null(city)){
    if (!is.character(city)){
      stop("city must be a character string")
    }
    url <- paste0(url, "&by_city=", str_replace_all(city, " ", "_"))
  }
  
  if (!is.null(type)){
    if (!is.character(type)){
      stop("type must be a character string")
    }
    type <- tolower(type)
    type_list <- c("micro", "nano", "regional", "brewpub", "large", "planning",
                    "bar", "contract", "proprietor", "closed")
    if (!(type %in% type_list)){
    stop("type must be either micro, nano, regional, brewpub, large, planning, bar, contract, proprietor, or closed")
    }
    url <- paste0(url, "&by_type=", str_replace_all(type, " ", "_"))
  }
  
  if (!is.null(name)){
    if (!is.character(name)){
      stop("brewery must be a character string")
    }
    url <- paste0(url, "&by_name=", name)
  }
  
  if(!is.null(state)){
    if (!is.character(state)){
      stop("state must be a character string")
    }
    if (usa){
      if (nchar(state) == 2){
        state <- toupper(state)
        switch(state,
                      AL =  state <- "Alabama",
                      AK =  state <- "Alaska",
                      AZ =  state <- "Arizona",
                      AR =  state <- "Arkansas",
                      CA =  state <- "California",
                      CO =  state <- "Colorado",
                      CT =  state <- "Connecticut",
                      DE =  state <- "Delaware",
                      DC =  state <- "District_of_Columbia",
                      FL =  state <- "Florida",
                      GA =  state <- "Georgia",
                      HI =  state <- "Hawaii",
                      ID =  state <- "Idaho",
                      IL =  state <- "Illinois",
                      IN =  state <- "Indiana",
                      IA =  state <- "Iowa",
                      KS =  state <- "Kansas",
                      KY =  state <- "Kentucky",
                      LA =  state <- "Louisiana",
                      ME =  state <- "Maine",
                      MD =  state <- "Maryland",
                      MA =  state <- "Massachusetts",
                      MI =  state <- "Michigan",
                      MN =  state <- "Minnesota",
                      MS =  state <- "Mississippi",
                      MO =  state <- "Missouri",
                      MT =  state <- "Montana",
                      NE =  state <- "Nebraska",
                      NV =  state <- "Nevada",
                      NH =  state <- "New_Hampshire",
                      NJ =  state <- "New_Jersey",
                      NM =  state <- "New_Mexico",
                      NY =  state <- "New_York",
                      NC =  state <- "North_Carolina",
                      ND =  state <- "North_Dakota",
                      OH =  state <- "Ohio",
                      OK =  state <- "Oklahoma",
                      OR =  state <- "Oregon",
                      PA =  state <- "Pennsylvania",
                      PR =  state <- "Puerto_Rico",
                      RI =  state <- "Rhode_Island",
                      SC =  state <- "South_Carolina",
                      SD =  state <- "South_Dakota",
                      TN =  state <- "Tennessee",
                      TX =  state <- "Texas",
                      UT =  state <- "Utah",
                      VT =  state <- "Vermont",
                      VA =  state <- "Virginia",
                      WA =  state <- "Washington",
                      WV =  state <- "West_Virginia",
                      WI =  state <- "Wisconsin",
                      WY =  state <- "Wyoming",
                      stop("usa state must be a 2 character abbreviation or name")
               )
      }
    }
    url <- paste0(url, "&by_state=", state)
  }
  
  if (!is.null(postal)){
    if ((!is.character(postal)) | 
        ((nchar(postal) != 5) & 
         (nchar(postal) != 10))){
      stop("Postal code must be a 5 or 10 character string in the form \"12345-6789\" or \"12345\"")
    }
    url <- paste0(url, "&by_postal=", postal)
  }
  
  if ((!is.null(latitude)) & (!is.null(longitude))){
    if ((!is.numeric(latitude)) | (!is.numeric(longitude))){
      stop("Latitude and longitude must be numeric values")
    }
    url <- paste0(url, "&by_dist=", latitude, ",", longitude)
  }

  df <- fromJSON(rawToChar(GET(url)$content))
  if (length(df) == 0 ){
    message("No results found. Try different parameters")
  }
  else {
    return(df)
  }
}
```
# Exploratory Data Analysis  

Now that we can access the endpoint with up to six different modifications, we can get some data
to analyze. Let's start by grabbing some data without any modifications using `get_brewery()`
```{r}
noMod <- get_brewery()
```

Given that most of the results are from the U.S., it seems reasonable to assume we should land
somewhere in Kansas (the center of the U.S.) using the means of the latitudes and longitudes 
if the results from the call are representative -- that is, they are unbiased and evenly spread.
We started by taking the unmodified data and subsetting it by selecting only the `name`, `city`,
`state`, `latitude`, and `longitude` variables. Really, the only necessary variables to perform
the calculations are `latitude` and `longitude`, but the others provide nice labels for
understanding the locations of the breweries. From there, we used `mutate()` to add new variables
to our dataset that find the means for both `latitude` and `longitude`, being careful to set 
them as numeric variables and ignoring `NA` values to ensure the calculation would work. Finally,
we created a `distFromMean` variable using the familiar distance formula 
d = sqrt[(x~2~ - x~1~)^2^ + (y~2~ - y~1~)^2^]

```{r}
usDist <- noMod %>%
  select(name, city, state, latitude, longitude) %>%
  mutate(meanLat = mean(as.numeric(latitude), na.rm = TRUE), 
         meanLong = mean(as.numeric(longitude), na.rm = TRUE),
         distFromMean = sqrt(((as.numeric(latitude) - meanLat)^2) +
           ((as.numeric(longitude)- meanLong)^2)))
usDist
```

How did we do? We can do another call using `get_brewery(state = "kansas")` to check.
Because we're only doing a quick check against the `latitude` and `longitude` values,
we're not concerned with grabbing the other variables this time. We can compare 
`meanLat` and `meanLong` with the latitude and longitude values from
breweries located in Kansas. As a side note, this API allows for searching by partial
match, so calling `get_brewery(state = "kansas")` returns results for both Kansas
and Arkansas.

```{r}
get_brewery(state = "kansas") %>%
  select(latitude, longitude)
```

We were close, but not quite what we were hoping for. Some knowledge of geography
lets us know we are just slightly north into the bordering state of Nebraska. But,
we can also perform another call using `get_brewery(state = "nebraska")` to verify.

```{r}
get_brewery(state = "nebraska") %>%
  select(latitude, longitude)
```

We can see that these values more closely align with the means we found. So what
happened? If we look back at `usDist`, we can see one glaring outlier from 
Killeshin, Ireland. Let's remove that and see if our results are better. Because
the mean values in `usDist` were calculated using all observations, we need to 
remember to start over with the unmodified data `noMod`. Again, we select all
variables of interest with `select()`, but this time we can use a `filter()`
function to select observations where `city` does not equal "Killeshin." Then,
we simply calculate the means as before with the subsetted data.


```{r}
noMod %>%
  select(name, city, state, latitude, longitude) %>%
  filter(city != "Killeshin") %>%
  mutate(meanLat = mean(as.numeric(latitude), na.rm = TRUE), 
         meanLong = mean(as.numeric(longitude), na.rm = TRUE),
         distFromMean = sqrt(((as.numeric(latitude) - meanLat)^2) +
           ((as.numeric(longitude)- meanLong)^2))) %>%
  arrange(state)
```

Now we are a bit further away than expected, so it appears that although close, this
data is not quite representative or unbiased. Visual inspection of the tibble shows
us that 6 results are from Oregon, with California and Colorado having 4 each; it
seems reasonable to assume this is why our means are northwest of where we expected.

A visual representation can help solidify our findings. We can use a histogram for this.
We started by creating a base plot object `g` using `ggplot()` with `usDist` as the input
and using an `aes()` statement to set the x-value to `distFromMean` by default.
Then we used `geom_histogram()` to create the actual histogram layer and setting the
outline color of the bars to blue and filling the bars with white. Finally, a `labs()`
statement was used to name the axes and title and the `theme()` statement was used
to center the title.

Here, the histogram includes the brewery from Ireland. Though clearly skewed right,
it looks like it could be considered nearly normally-distributed without the outlier.
The x-axis shows us the distance from the mean and each bar's height represents the
frequency that breweries were that corresponding distance from the mean.

```{r}
g <- ggplot(usDist, aes(x = distFromMean))
g + geom_histogram(color = 4, fill = "white") +
  labs(title = "Distance of Brewery from Mean", x = "Distance", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

Here is the histogram without the brewery from Ireland. Notice we used `na.omit()`
to omit the `NA` values that showed up for `latitude` and `longitude` (which can't
be properly plotted anyway), but because "Killeshin" is the only city outside of 
the U.S., it was a convenient way to exclude it from the data. Without the outlier,
the plot shows a lot of gaps. Because this is representing continuous data, we felt
it was appropriate to experiment with `binwidth =` and `bins =` to minimize the gaps.
Ultimately, we settled on `bins = 10`, while keeping the color, fill, and labels the
same as before. 

As with the previous plot, this histogram looks nearly normally-distributed. This 
indicates that it is nearly representative of the population.
Still, it is important to note that we couldn't make assumptions about the population
with this data. As with the above histogram, the x-axis is the distance from the mean and 
the y-axis consists of the corresponding frequencies of breweries' distances from the mean.

```{r}
g <- ggplot(na.omit(usDist), aes(x = distFromMean))
g + geom_histogram(color = 4, fill = "white", bins = 10) +
  labs(title = "Distance of Brewery from Mean", x = "Distance", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```


Next, let's look at some more specific data by calling the function `get_brewery(state = "ny")`
to get brewery data for the state of New York.
```{r}
#Get brewery data for NY
ny <- get_brewery(state = "ny")
```

We want to see if there are any patterns by type of brewery. To do this, we will investigate
latitude and longitude values across brewery types. Additionally, we may want to look at
the means of each type to see if they have significant variance.

We started by creating a new object called `nyDist` that uses the modified `ny` dataset
and selects the `name`, `brewery_type`, `latitude`, and `longitude` variables. Because
we're interested in whether patterns exist by brewery type, we used `group_by(brewery_type)`
to group the data by `brewery_type` and perform separate calculations per level of 
the groups. We used `mutate()` again to add columns for newly calculated `meanLat`
and `meanLong`; again we had to remember to set the variables to numeric using
`as.numeric` and being careful to ignore `NA` values by setting `na.rm = TRUE`.
Finally, we used the `arrange()` function to arrange the data by our newly created
variable.

```{r}
nyDist <- ny %>% 
  select(name, brewery_type, latitude, longitude) %>%
  group_by(brewery_type) %>%
  mutate(meanLat = mean(as.numeric(latitude), na.rm = TRUE), 
         meanLong = mean(as.numeric(longitude), na.rm = TRUE)) %>%
  arrange(meanLat)
nyDist
```

Now that we have the data, let's create a scatter plot to look for obvious patterns.
As before, we created a base `g` object using `ggplot()` and we used `na.omit()`
again -- this time to make sure `NA` doesn't show up on the y-axis. This time we
added a `y =` to the `aes()` statement since a scatter plot needs a variable on 
both axes. We used a `geom_point()` layer to create the actual scatter plot and 
specified `color = brewery_type` in the `aes()` statement to display breweries
by type using different colors (and automatically creating a nice legend). As before,
a `labs()` layer was used to name the axes and give the plot a title. The first `theme()`
layer was used to rotate the x-axis 90 degrees so the longitude values could be 
legible, while the second `theme()` layer was used to center the title.

```{r}
g <- ggplot(na.omit(nyDist), aes(x = longitude, y = latitude))
g + geom_point(aes(color = brewery_type)) +
    labs(color = "Brewery Type", x = "Longitude", y = "Latitude") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
    labs(title = "NY Breweries by Type", hjust = 0.5) +
    theme(plot.title = element_text(hjust = 0.5))
```

As the labels would suggest, this plots the longitude vs. latitude of each brewery
in our dataset. Looking at the plot, there doesn't appear to be any sort of obvious
pattern that would suggest a preference of location to build a brewery, but there
are clusters or groups of breweries by type. We believe this suggests the influence
of competition. 

Finally, let's look at a breakdown of some summary statistics for this numerical data
at each setting of the `brewery_type` variable. To do this, we used the same `nyDist`
dataset in conjunction with the `summarize()` function to print a tibble of `meanLat`,
`meanLong`, along with `sdLat` and `sdLong` that we created using the built-in `sd()`
function. Like in previous calculations, it was important to set `latitude` and 
`longitude` as numeric values using `as.numeric()` and omitting `NA` values from
the calculations using `na.rm = TRUE`.

The values of `meanLat` and `meanLong` are very similar across all categories of
`brewery_type`. We believe this suggests an even spattering of types across all
available locations -- this could be understood to indicate that no major confounders
exist that dictate where a type of brewery may be established.

```{r}
nyDist %>%
  summarize(meanLat = mean(as.numeric(latitude), na.rm = TRUE), 
            meanLong = mean(as.numeric(longitude), na.rm = TRUE),
            sdLat = sd(latitude, na.rm = TRUE), sdLong = sd(longitude, na.rm = TRUE))
```


Derive Region variable from state, make contingency tables, create bar graph.

```{r}
brew <- get_brewery()

brew_reg <- brew %>%
  mutate(Region = 
    if_else(brew$state %in%
    c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", 
      "Vermont", "New Jersey", "New York", "Pennsylvania"),
    "Northeast",
    if_else(brew$state %in%
    c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", 
      "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota"),
    "Midwest",
    if_else(brew$state %in%
    c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina",
      "South Carolina", "Virginia", "Washington D.C.", "West Virginia", 
      "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas",
      "Louisiana", "Oklahoma", "Texas"),
    "South",
    if_else(brew$state %in%
    c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah",
      "Wyoming", "Alaska", "California","Hawaii","Oregon","Washington"),
    "West", 
    "Not US"))))) %>%
  select(id:state, Region, everything())

table(brew_reg$Region)
table(brew_reg$brewery_type)
table(brew_reg$Region, brew_reg$brewery_type)

g <- ggplot(brew_reg, aes(x = Region))
g + geom_bar(aes(fill = brewery_type)) +
    labs(x = "Geographic Region",
         y = "Number of Breweries",
         title = "Bar Plot of Breweries by US Region") + 
    scale_fill_discrete(name = "Brewery Type", 
                        labels = c("Brewpub", "Closed", "Contract", "Large",
                                   "Micro", "Proprietor")
                        )

micro <- get_brewery(type = "micro")
large <- get_brewery(type = "large")
combined <- bind_rows(micro, large)

g <- ggplot(combined, aes(x = brewery_type, y = as.numeric(latitude)))
g + geom_boxplot(fill = "red", na.rm = TRUE) +
    geom_jitter() +
    labs(title = "Boxplot of Latitudes by Brewery Type",
         x = "Brewery Type",
         y = "Latitude in Degrees")

g <- ggplot(combined, aes(x = brewery_type, y = as.numeric(longitude)))
g + geom_boxplot(fill = "blue", na.rm = TRUE) +
    geom_jitter() +
    coord_flip() +
    labs(title = "Boxplot of Longitudes by Brewery Type",
         x = "Brewery Type",
         y = "Longitude in Degrees")

```
